Interactive Curvature Tensor Visualization on Digital Surfaces\thanks{This work has been mainly funded by XXXXXXX research grants.}

### 2015-10-21 latex/article.tex
wip

\sloppy In our context, we consider digital shapes (any subset of $\Z^d$) and
boundaries of digital shapes. We denote by $\DSh$ the Gauss
digitization of $\Shape$ in a $d-$dimensional grid with grid step $h$,
\emph{i.e.}  $\DSh = \Shape \cap (h\Z)^d$. For such digitized set
$\DigShape\EqDef\DSh$, $\Body{\DigShape}{h}$ is a subset of $\R^d$ corresponding
to the union of hypercubes centered at $h\cdot \vz$ for  $\vz\in Z$
with edge length $h$. By doing so, both $\partial\Shape$ and
$\partial\Body{Z}{h}$ are topological boundary of objects lying in the
same space (See
Fig. \ref{fig:notations}-b). Note that the combinatorial digital boundary
of $Z$ made of cells in a cellular Cartesian complex (\emph{pointels},
\emph{linels}, {\em surfels}, \ldots), can be trivally embedded into $\R^d$
such that it coincides with $\partial\Body{Z}{h}$.

\begin{figure}[t]{\small
    \begin{center}
      \subfigure[]{\begin{overpic}[width=5cm]{figs/integral2D_2}
          \put(42,28){$\times$}
          \put(46,31){$\vx$}
          \put(55,45){$B_R(\vx)$}
          \put(1,10){$\dS$}
          \put(20,10){$\Shape$}
      \end{overpic}}
      \subfigure[]{\begin{overpic}[width=5cm]{figs/notation_2}
          \put(44.5,25){$\vp$}
          \put(55,45){$B_R(\vx)$}
          \put(42,28.5){$\times$}
          \put(45,32.5){$\vx$}
          \put(45,16.5){$h$}
          %\put(88,39){$\Proj(p)$}
          \put(3,11){$\partial \Shape$}
          \put(15,7){$\partial \Body{\DigShape}{h}$}
      \end{overpic}}
      \vspace{-0.4cm}
    \end{center}}
    \caption{Integral invariant computation $(a)$ and
      notations $(b)$ in dimension 2 \cite{CVIU2014}.\label{fig:notations}}
\end{figure}

In \cite{DGCI2013}, we have defined the 2D digital curvature estimator
$\CurvH{R}$ and the 3D digital mean curvature estimator $\MeanCurvH{R}$ based on
the digital volumetric integral $V_{R_d}$ (or $A_{R_d}$ in dimension 2) of the
intersection between the shape $\Shape$ and a ball of digital radius $R_d$:
%
\begin{Definition}
  %
  Given the Gauss digitization $\Body{\DSh}{h}$ of a shape $\Shape \subset \R^2$
  (or $\R^3$ for the 3D mean curvature estimator) scaled by the digitization
  step $h$, {\em digital curvature estimators} are defined for any point $\vx
  \in \R^2$ (or $\R^3$) as:
  %
  \begin{align}
    \forall 0 < h < R,\quad
    &\CurvH{R}(\Body{\DSh}{h},\vx,h) \EqDef \frac{3\pi}{2R}
    - \frac{3\AreaC(\Ball{R/h}{\vx/h} \cap
    \Body{\DSh}{h}, h)}{R^3}\,,\label{eq:dig-curvature-estimator-k}\\
    &\MeanCurvH{R}(\Body{\DSh}{h},\vx,h) \EqDef \frac{8}{3R}
    - \frac{4\VolC(\Ball{R/h}{\vx/h} \cap
    \Body{\DSh}{h}, h)}{\pi R^4}\,.\label{eq:dig-curvature-estimator-H}
  \end{align}
  %
  where $\AreaC(\Ball{R/h}{\vx/h} \cap \Body{\DSh}{h}, h) = h^2 \MCard(
  \Ball{R/h}{\vx/h} \cap \Body{\DSh}{h} )$ and $\VolC(\Ball{R/h}{\vx/h} \cap
  \Body{\DSh}{h}, h) = h^3 \MCard( \Ball{R/h}{\vx/h} \cap \Body{\DSh}{h} )$ are
  the number of digital points in the intersection between $\Body{\DSh}{h}$ and
  the ball $\Ball{R/h}{\vx/h}$ of radius $R/h$ centered on $\vx/h$, scaled by
  $h^d$.\todo[inline]{J: Hum... pas top...}
  %
  % Similarly, $\hat{\lambda}_1$
  % and $\hat{\lambda}_2$ are the two greatest eigenvalues of the covariance
  % matrix of the digital points in the intersection between $Z$ and the ball.
  %
\end{Definition}
%
We have proved that estimating the 2D curvature and 3D mean curvature on digital
objets with volumetric integral information have the multigrid convergence
property (setting $R = kh^\frac{1}{3}$ on convex shapes, where $k$ is a constant
related to the maximal curvature of the shape, with at least $C^3$-boundary and
bounded curvature) \cite{DGCI2013}. For short, when the digital objet becomes
finer and finer, \ie when the digitization step $h$ tends to zero, the estimated
quantity on $\Bd{\Body{\DSh}{h}}$  converges (theoretically and experimentally)
to the associated one on $\Bd{\Shape}$. More precisely, we have:
%
\begin{align}
\forall 0 < h \le h_0,\,\, &\forall \vx \in \Bd{\Shape},\,\,
\forall \vxH \in \Bd{\Body{\DSh}{h}} \text{~with~} \| \vxH -\vx\|_\infty \le h, \nonumber \\
&\CurvH{R}(\DSh, \vxH, h) = \Curv(\Shape, \vx) + O\left(h^{\frac{1}{3}}\right)\,,\\
&\MeanCurvH{R}(\DSh, \vxH, h) = \MeanCurv(\Shape, \vx) + O\left(h^{\frac{1}{3}}\right)\,.
\end{align}

@@

We have shown that the radius of the ball depends to the geometry of the
underlying shape. In \cite{DGCI2014}, we proposed a parameter-free estimation of
the radius of the ball by analyzing the shape \wrt the local geometry through to
the maximal digital straight segments of the boundary of the digital shape. In
\cite{SMI2015}, we analyze in scale-space (\ie with a range of radii) curvature
estimation of our algorithm for a given digital shape. This allows us to detect
features of the shape thanks to the behavior on singularities. In both methods, we need to explore the radius of the ball at different sizes and the computation time, depending to the size of the digital objet and the radius of the ball, is an obstacle.
\todo[inline]{Faut faire une transition...}
Computing curvatures from our algorithm in real-time can be really useful.

@@

Once decisions have been made, a new set of cells are sent to the mesh
generation step described above. Updating the cell buffer can be
efficiently implemented on GPU hardware as illustrated in
Figure~\ref{fig_lod_octree}-\emph{right}.

Finally, before constructing the triangulation from remaining cells
and transition cells, geometrical culling is performed in order to
skip the triangle patch construction which are not visible. Figure
\ref{fig_pipeline} illustrates the overall fully data parallel
pipeline.

@@

\section{Interactive Computation on GPU}
\label{sec:inter-visu-gpu}

\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{figs/approx}
  \end{center}
  \caption{BALBLABLABLA.}
  \label{fig:approx}
\end{figure}

>>>>

\sloppy In our context, we consider digital shapes (any subset of
$\Z^d$) and boundaries of digital shapes. We denote by $\DSh$ the
Gauss digitization of $\Shape$ in a $d-$dimensional grid with grid
step $h$, \emph{i.e.}  $\DSh \EqDef \{ \vz \in \Z^d,\, h\cdot\vz\in\Shape\}$. For such
digitized set $\DigShape\EqDef\DSh$, $\Body{\DigShape}{h}$ is a subset
of $\R^d$ corresponding to the union of hypercubes centered at $h\cdot
\vz$ for $\vz\in Z$ with edge length $h$. By doing so, both
$\partial\Shape$ and $\partial\Body{Z}{h}$ are topological boundary of
objects lying in the same space (See Fig. \ref{fig:notations}-b). Note
that the combinatorial digital boundary of $Z$ made of cells in a
cellular Cartesian complex (\emph{pointels}, \emph{linels}, {\em
  surfels}, \ldots), can be trivally embedded into $\R^d$ such that it
coincides with $\partial\Body{Z}{h}$.

\begin{figure}[t]{\small
    \begin{center}
      \subfigure[]{\begin{overpic}[width=5cm]{figs/integral2D_2}
          \put(42,28){$\times$}
          \put(46,31){$\vx$}
          \put(55,45){$B_R(\vx)$}
          \put(1,10){$\dS$}
          \put(20,10){$\Shape$}
      \end{overpic}}~~~~~~
      \subfigure[]{\begin{overpic}[width=5cm]{figs/notation_2}
          \put(44.5,25){$\vp$}
          \put(55,45){$B_R(\vx)$}
          \put(42,28.5){$\times$}
          \put(45,32.5){$\vx$}
          \put(45,16.5){$h$}
          %\put(88,39){$\Proj(p)$}
          \put(3,11){$\partial \Shape$}
          \put(15,7){$\partial \Body{\DigShape}{h}$}
      \end{overpic}}
      \vspace{-0.4cm}
    \end{center}}
    \caption{Integral invariant computation $(a)$ and
      notations $(b)$ in dimension 2 \cite{CVIU2014}.\label{fig:notations}}
\end{figure}

In \cite{DGCI2013}, we have defined the 2D digital curvature estimator
$\CurvH{R}$ and the 3D digital mean curvature estimator $\MeanCurvH{R}$ based on
the digital volumetric integral $V_{R_d}$ (or $A_{R_d}$ in dimension 2) of the
intersection between the shape $\Shape$ and a ball of digital radius $R_d$:
%
\begin{Definition}
  %
  Given the Gauss digitization $Z\EqDef\DSh$ of a shape $\Shape \subset \R^2$
  (or $\R^3$ for the 3D mean curvature estimator), {\em digital curvature estimators} are defined for any point $\vx
  \in \R^2$ (or $\R^3$) as:
  %
  \begin{align}
    \forall 0 < h < R,\quad
    &\CurvH{R}(Z,\vx,h) \EqDef \frac{3\pi}{2R}
    - \frac{3\AreaC(\Ball{R/h}{\vx/h} \cap
    {Z}, h)}{R^3}\,,\label{eq:dig-curvature-estimator-k}\\
    &\MeanCurvH{R}(Z,\vx,h) \EqDef \frac{8}{3R}
    - \frac{4\VolC(\Ball{R/h}{\vx/h} \cap
    Z, h)}{\pi R^4}\,.\label{eq:dig-curvature-estimator-H}
  \end{align}
  %
  where $\AreaC(\Ball{R/h}{\vx/h} \cap \Body{\DSh}{h}, h) \EqDef h^2 \MCard(
  \Ball{R/h}{\vx/h} \cap \Body{\DSh}{h} )$ and $\VolC(\Ball{R/h}{\vx/h} \cap
  \Body{\DSh}{h}, h) \EqDef h^3 \MCard( \Ball{R/h}{\vx/h} \cap \Body{\DSh}{h} )$ are
  the number of digital points in the intersection between $\Body{\DSh}{h}$ and
  the ball $\Ball{R/h}{\vx/h}$ of radius $R/h$ centered on $\vx/h$, scaled by
  $h^d$.\todo[inline]{J: Hum... pas top...}
  %
  % Similarly, $\hat{\lambda}_1$
  % and $\hat{\lambda}_2$ are the two greatest eigenvalues of the covariance
  % matrix of the digital points in the intersection between $Z$ and the ball.
  %
\end{Definition}
%
We have proved that estimating the 2D curvature and 3D mean curvature on digital
objets with volumetric integral information have the multigrid convergence
property (setting $R = kh^\frac{1}{3}$ on convex shapes, where $k$ is a constant
related to the maximal curvature of the shape, with at least $C^3$-boundary and
bounded curvature) \cite{DGCI2013}. For short, when the digital objet becomes
finer and finer, \ie when the digitization step $h$ tends to zero, the estimated
quantity on $\Bd{\Body{\DSh}{h}}$  converges (theoretically and experimentally)
to the associated one on $\Bd{\Shape}$. More precisely, we have:
%
\begin{align}
\forall 0 < h \le h_0,\,\, &\forall \vx \in \Bd{\Shape},\,\,
\forall \vxH \in \Bd{\Body{\DSh}{h}} \text{~with~} \| \vxH -\vx\|_\infty \le h, \nonumber \\
&\CurvH{R}(\DSh, \vxH, h) = \Curv(\Shape, \vx) + O\left(h^{\frac{1}{3}}\right)\,,\\
&\MeanCurvH{R}(\DSh, \vxH, h) = \MeanCurv(\Shape, \vx) + O\left(h^{\frac{1}{3}}\right)\,.
\end{align}

@@

We have shown that the radius of the ball depends on the geometry of
the underlying shape. In \cite{DGCI2014}, we proposed a parameter-free
estimation of the radius of the ball by analyzing the shape \wrt the
local shape geometry using maximal digital straight segments of the
digital boundary. In \cite{SMI2015}, we analyze in scale-space (\ie
for a range of radii) curvature estimation of our algorithm for a
given digital shape. This allows us to detect features of the shape
thanks to the behavior on singularities.

As a consequence, for all these integral invariant based approaches,
we need to consider different ball radius which could be time
consuming when implemented on CPU. We propose here a fully parallel
implementation on GPU allowing us to change the radius and thus update
the estimated quantities in realtime.

@@

Once decisions have been made, a new set of cells are sent to the mesh
generation step described above. Updating the cell buffer can be
efficiently implemented on GPU hardware as illustrated in
Figure~\ref{fig_lod_octree}-\emph{right}.  Finally, before
constructing the triangulation from remaining cells and transition
cells, geometrical culling is performed in order to skip the triangle
patch construction which are not visible. Figure \ref{fig_pipeline}
illustrates the overall fully data parallel pipeline.

@@

\section{Interactive Curvature Computation on GPU}
\label{sec:inter-visu-gpu}

\subsection{Design principle}

The general idea is to perform and Integral Invariant computation on
GPU at each vertex of the generated triangular mesh. First note that
of the LoD criterion is removed, each MC vertex is exactly centered at
a surfel center. At different depth of the octree, MC vertices still
correspond to surfel centers of subsampled versions of the input
object. Hence, Integral Invariant framework defined in
Section~\ref{sec:curv-tens-estim} is consistent with the triangular
surface obtained on GPU: triangles are used for visualization purposes
but all computations are performed on the digital object $Z$ and thus
estimators
(\ref{eq:dig-curvature-estimator-k}),
(\ref{eq:dig-curvature-estimator-H}), (\ref{eq:dig-curvature-estimator-k1})
and (\ref{eq:dig-curvature-estimator-k2}).

To implement the integration on $B_R(x)\cap\Shape$
(Fig.~\ref{fig:approx}-$a$), a naive approach consists in scanning all
digital points inside such domain (Fig.~\ref{fig:approx}-$b$) and then
to estimate the geometrical moment as the sum of the geometrical
moments of each elementary cubes lying in the intersection. On GPU,

\begin{figure}
  \begin{center}
    \subfigure[]{\includegraphics[width=5cm]{figs/approx1}}
    \subfigure[]{\includegraphics[width=5cm]{figs/approx-reg}}
    \subfigure[]{\includegraphics[width=5cm]{figs/approx-reg-2}}
    \subfigure[]{\includegraphics[width=5cm]{figs/approx-bh}}
    \subfigure[]{\includegraphics[width=5cm]{figs/approx-ada}}
  \end{center}
  \caption{BALBLABLABLA.}
  \label{fig:approx}
\end{figure}

\subsection{Realtime computation on GPU}
